Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-02-28 paper/main.tex
refined discussion of fig 6 results

For the initial quizzes participants took (prior to watching either lecture),
participants' estimated knowledge tended to be low overall, and relatively
unstructured (Fig.~\ref{fig:predictions}, left panel). When we held out
individual questions and estimated their knowledge at the held-out questions'
embedding coordinates, we found no reliable differences in the estimates when
the held-out question had been correctly versus incorrectly answered ($t(633) =
0.577,~p = 0.564$). After watching the first video, estimated knowledge for
held-out correctly answered questions (from the second quiz;
Fig.~\ref{fig:predictions}, middle panel) exhibited a positive shift relative
to held-out incorrectly answered questions ($t(633) = 3.961,~p < 0.001$). After
watching the second video, estimated knowledge (from the third quiz;
Fig.~\ref{fig:predictions}, right panel) for \textit{all} questions exhibited a
positive shift. However, the increase in estimated knowledge for held-out
correctly answered questions was larger than for held-out incorrectly answered
questions (estimated knowledge for correctly versus incorrectly answered
Quiz~3 questions: $t(628) = 2.045,~p = 0.041$).

>>>>

For the initial quizzes participants took (prior to watching either lecture),
participants' estimated knowledge tended to be low overall, and relatively
unstructured (Fig.~\ref{fig:predictions}, left panel). When we held out
individual questions and estimated their knowledge at the held-out questions'
embedding coordinates, we found no reliable differences in the estimates when
the held-out question had been correctly versus incorrectly answered ($t(633) =
0.577,~p = 0.564$). This reflects a floor effect: when knowledge is low
everywhere, there is little signal to differentiate between what is known
versus unknown. After watching the first lecture, estimated knowledge for
held-out correctly answered questions (from the second quiz;
Fig.~\ref{fig:predictions}, middle panel) exhibited a positive shift relative
to held-out incorrectly answered questions ($t(633) = 3.961,~p < 0.001$). This
second quiz provides the maximally sensitive test for our knowledge
predictions, since (if knowledge is estimated accurately) participants' Quiz~2
responses should demonstrate specific knowledge about Lecture~1 content, but
knowledge about Lecture~2 and general physics concepts should be roughly
unchanged from before they watched Lecture~1. After watching the second lecture,
estimated knowledge (from the third quiz; Fig.~\ref{fig:predictions}, right
panel) for \textit{all} questions exhibited a positive shift. However, the
estimated knowledge for held-out correctly answered questions remained
greater than that for held-out incorrectly answered questions ($t(628) = 2.045,~p
= 0.041$). This third contrast reflects a ceiling effect: when knowledge is
relatively high everywhere, the signal differentiating what is known versus
unknown is relatively weak. Taken together, this set of analyses demonstrates
that our knowledge prediction framework is most informative when participants
exhibit variability in their knowledge of the content captured by the text
embedding model.
